---
title: "FishersTest"
author: "Anna Lewkowicz"
date: "2023-07-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Setup 

Load the necessary library

```{r libraries, warning=FALSE, message=FALSE}
library(dplyr) 
```

## Read all the files

```{r read}
combinations_counts <- read.csv("data/countedPairs/counted_pairs.csv", head = TRUE)
```

Check for missing values

```{r checkNA}
sum(is.na(combinations_counts))
```

Remove all rows with NA values

```{r removeNA}
combinations_counts <- na.omit(combinations_counts)
```

Remove rows with zeros in columns with co-occuring mutations and CNVs
Create separate dataframes for positive and negative CNVs

```{r separateDF}
positive <- combinations_counts %>%
  filter(cnv_positive_mut > 0)
negative <- combinations_counts %>%
  filter(cnv_negative_mut > 0)
```

## Perform Fisher's exact test for each gene-cancer pairs

```{r fisher}
for (i in 1:nrow(positive)) {
  matrix_values <- matrix(c(positive[i, "cnv_positive_mut"], positive[i, "cnv_positive"], 
                            positive[i, "mut_only"], positive[i, "nothing"]),
                          nrow = 2)
  
  fisher_result <- fisher.test(matrix_values)
  positive[i, "p.value"] <- fisher_result$p.value
  positive[i, "odds"] <- fisher_result$estimate

}

for (i in 1:nrow(negative)) {
  matrix_values <- matrix(c(negative[i, "cnv_negative_mut"], negative[i, "cnv_negative"], 
                            negative[i, "mut_only"], negative[i, "nothing"]),
                          nrow = 2)
  
  fisher_result <- fisher.test(matrix_values)
  negative[i, "p.value"] <- fisher_result$p.value
  negative[i, "odds"] <- fisher_result$estimate

}
```

Check number of p.values below 0.1

```{r checkP}
sum(positive$p.value < 0.1)
sum(negative$p.value < 0.1)
```

Adjust p-values

```{r adjpValues}
positive$p.adj <- p.adjust(positive$p.value, method = "BH")
negative$p.adj <- p.adjust(negative$p.value, method = "BH")
```

Check number of p.adj.values below 0.1

```{r checkAdj}
sum(positive$p.adj < 0.1)
sum(negative$p.adj < 0.1)
```

Save files

```{r save, warning=FALSE}
dir.create("data/fishersTest", recursive = TRUE)
write.csv(positive, "data/fishersTest/fishers_positive.csv", row.names = FALSE)
write.csv(negative, "data/fishersTest/fishers_negative.csv", row.names = FALSE)
```