---
title: "MakeGraphs"
author: "Anna Lewkowicz"
date: "2023-07-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup 

Load the necessary library

```{r libraries, warning=FALSE, message=FALSE}
library(ggplot2)
```

## Load the data

```{r loadData}
combinations_counts = read.csv("data/countedPairs/combinations_counts.csv", head = TRUE)
```

## Extract the data of interest for "all data" plot 

Remove NA values 

```{r removeNA}
combinations_counts_clean <- combinations_counts[!is.na(combinations_counts$acronym), ]
```

Find top 10 gene cancer pairs

```{r top10}
top10 <- head(combinations_counts_clean[
  order(combinations_counts_clean$cnv_positive_mut + 
          combinations_counts_clean$cnv_negative_mut, 
        decreasing = TRUE),], 10)
```

Transform top_10 into data frame

```{r transformToDF}
top_10 <- data.frame(top10)
```

Merge the first two columns with a vertical bar separator

```{r mergeTwoColumns}
top_10$type <- paste(top_10[,1], top_10[,2], sep = " | ")
```

Remove the original columns

```{r removeOriginalColumns}
top_10 <- top_10[, -c(1, 2)]
```

Move the merged_columns column to the front

```{r moveMergedColumns}
top_10 <- top_10[, c(ncol(top_10), 1:(ncol(top_10)-1))]
```

Calculate the proportion of each value in a row

```{r calculateProportions}
proportions_df <- round(as.data.frame(t(apply(top_10[,2:7], 1, function(row) row / sum(row)))), 4)
```

Add type column to the proportion dataframe

```{r addColumn}
rownames(proportions_df) <- top_10$type
```

First, convert the row names into a separate column of the dataframe

```{r converRowNames}
proportions_df$gene <- rownames(proportions_df)
rownames(proportions_df) <- NULL
```

Then, reshape the dataframe into a long format 

```{r reshape}
proportions_long <- tidyr::gather(proportions_df, key = "category", value = "proportion", -gene)

```

## Create "all data" plot 

Create the plot using ggplot

```{r ggp1, fig.align = "center"}
ggp1 <- ggplot(proportions_long, aes(x = gene, y = proportion, fill = category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "All data", x = "Gene", y = "Proportions") +
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00")) +
  theme_classic() +
  theme(legend.position = "bottom",
    axis.text.x = element_text(angle = 45, size = 8, hjust = 1))

ggp1
```

## Save "all data" plot 

Save the plot as a PNG file

```{r ggp1Save, warning=FALSE, message=FALSE}
ggsave("data/graphics/All_data.png", plot = ggp1, dpi = 300)
```

## Extract the data of interest for "Proportions by Gene" plot 
### REMOVE "NOTHING" COLUMN
Remove 'nothing' column

```{r removeNothing}
top_10_only_mc <- top_10[,c(1:4,6:7)]
```

Calculate the proportion of each value in a row

```{r calculateProportions2}
proportions_only_mc <- round(as.data.frame(t(apply(top_10_only_mc[,2:6], 1, function(row) row / sum(row)))), 4)
```

Add type column to the proportion dataframe

```{r addTypeCol2}
rownames(proportions_only_mc) <- top_10_only_mc$type
```

First, convert the row names into a separate column of the dataframe

```{r convertRowNames2}
proportions_only_mc$gene <- rownames(proportions_only_mc)
rownames(proportions_only_mc) <- NULL
```

Then, reshape the dataframe into a long format 

```{r reshape2}
proportions_only_mc_long <- tidyr::gather(proportions_only_mc, key = "category", value = "proportion", -gene)
```

## Create "Proportions by Gene" plot 

Create the plot using ggplot

```{r ggp2, fig.align = "center"}
ggp2 <- ggplot(proportions_only_mc_long, aes(x = gene, y = proportion, fill = category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Proportions by Gene", x = "Gene", y = "Proportions") +
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")) +
  theme_classic() +
  theme(legend.position = "bottom",
    axis.text.x = element_text(angle = 45, size = 8, hjust = 1))

ggp2
```

## Save "Proportions by Gene" plot 

Save the plot as a PNG file

```{r ggp2Save, warning=FALSE, message=FALSE}
ggsave("data/graphics/Nothing_removed_plot.png", plot = ggp2, dpi = 300)
```

## Extract the data of interest for log transformed "Proportions by Gene" plot 
### LOG TRANSFORM 

```{r logTransform}
top_10_only_mc_log <- top_10_only_mc

top_10_only_mc_log[,c(2:6)] <- log(top_10_only_mc_log[,c(2:6)])

top_10_only_mc_log <- rapply(top_10_only_mc_log, function(x) ifelse(is.infinite(x), 0, x), how = "replace")
```

Calculate the proportion of each value in a row

```{r calculateProportions3}
proportions_only_mc_log <- round(as.data.frame(t(apply(top_10_only_mc_log[,2:6], 1, function(row) row / sum(row)))), 4)
```

Add type column to the proportion dataframe

```{r addColumn3}
rownames(proportions_only_mc_log) <- top_10_only_mc_log$type
```

First, convert the row names into a separate column of the dataframe

```{r convertRowNames3}
proportions_only_mc_log$gene <- rownames(proportions_only_mc_log)
rownames(proportions_only_mc_log) <- NULL
```

Then, reshape the dataframe into a long format 

```{r reshape3}
proportions_only_mc_long_log <- tidyr::gather(proportions_only_mc_log, key = "category", value = "proportion", -gene)
```

## Create log transformed "Proportions by Gene" plot 

Create the plot using ggplot

```{r ggp3, fig.align = "center"}
ggp3 <- ggplot(proportions_only_mc_long_log, aes(x = gene, y = proportion, fill = category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Proportions by Gene (log transformed)", x = "Gene", y = "Proportions") +
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")) +
  theme_classic() +
  theme(legend.position = "bottom",
    axis.text.x = element_text(angle = 45, size = 8, hjust = 1))

ggp3
```

## Save log transformed "Proportions by Gene" plot 

Save the plot as a PNG file

```{r ggp3Save, warning=FALSE, message=FALSE}
ggsave("data/graphics/Log_transformed_original.png", plot = ggp3, dpi = 300)
```

## Extract the data of interest for merged "Proportions by Gene" plot 

Merge cnv and cnv_mut

```{r mergeMutCNV}
top_10_only_mc_log_merged <- top_10_only_mc_log

top_10_only_mc_log_merged$cnv <- top_10_only_mc_log_merged$cnv_positive + top_10_only_mc_log_merged$cnv_negative

top_10_only_mc_log_merged$cnv_mut <- top_10_only_mc_log_merged$cnv_positive_mut + top_10_only_mc_log_merged$cnv_negative_mut

top_10_only_mc_log_merged <- top_10_only_mc_log_merged[,c(1,2,7,8)]
```

Calculate the proportion of each value in a row

```{r calcPropPerRow}
proportions_only_mc_log_shorten <- round(as.data.frame(t(apply(top_10_only_mc_log_merged[,2:4], 1, function(row) row / sum(row)))), 4)
```

Add type column to the proportion dataframe

```{r addTypeCol}
rownames(proportions_only_mc_log_shorten) <- top_10_only_mc_log_merged$type
```

First, convert the row names into a separate column of the dataframe

```{r convertRowNames}
proportions_only_mc_log_shorten$gene <- rownames(proportions_only_mc_log_shorten)
rownames(proportions_only_mc_log_shorten) <- NULL
```

Then, reshape the dataframe into a long format

```{r reshape4}
proportions_only_mc_long_log_shorten <- tidyr::gather(proportions_only_mc_log_shorten, key = "category", value = "proportion", -gene)
```

Reorder the levels of the category variable, so it looks better on the graph

```{r reorder}
proportions_only_mc_long_log_shorten$category <- factor(
  proportions_only_mc_long_log_shorten$category,
  levels = c("cnv_mut", "cnv", "mut_only")
)
```

## Create merged "Proportions by Gene" plot 

Create the plot using ggplot

```{r ggp4, fig.align = "center"}
ggp4 <- ggplot(proportions_only_mc_long_log_shorten, aes(x = gene, y = proportion, fill = category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Proportions by Gene (log transformed)", x = "Gene", y = "Proportions") +
  scale_fill_manual(values = c("#009E73", "#E69F00", "#0072B2")) +
  theme_classic() +
  theme(legend.position = "bottom",
    axis.text.x = element_text(angle = 45, size = 8, hjust = 1))

ggp4
```

## Save merged "Proportions by Gene" plot 

Save the plot as a PNG file

```{r ggp4Save, warning=FALSE, message=FALSE}
ggsave("data/graphics/Log_transformed_merged.png", plot = ggp4, dpi = 300)
```

## Volcano plots 

Load the data

```{r loadFishersNegative}
fishers_negative <- read.csv("data/fishersTest/fishers_negative.csv", head = TRUE)
```

Create volcano negative plot

```{r volcanoNegative}
volcano_plot_negative <- ggplot(fishers_negative, aes(x = odds, y = p.adj)) +
  geom_point(color = "blue", size = 3) +
  labs(title = "Volcano Plot",
       x = "Odds",
       y = "Adjusted p-value")
```

Save the plot

```{r volcanoNegativeSave, warning=FALSE, message=FALSE}
ggsave("data/graphics/volcano_fishers_negative.png", plot = volcano_plot_negative, dpi = 300)
```

Load the data

```{r loadFishersPositive}
fishers_positive <- read.csv("data/fishersTest/fishers_positive.csv", head = TRUE)
```

Create volcano positive plot

```{r volcanoPositive}
volcano_plot_positive <- ggplot(fishers_positive, aes(x = odds, y = p.adj)) +
  geom_point(color = "blue", size = 3) +
  labs(title = "Volcano Plot",
       x = "Odds",
       y = "Adjusted p-value")
```

Save the plot

```{r volcanoPositiveSave, warning=FALSE, message=FALSE}
ggsave("data/graphics/volcano_fishers_positive.png", plot = volcano_plot_positive, dpi = 300)
```
