---
title: "Preprocess"
author: "Anna Lewkowicz"
date: "2023-07-11"
output: html_document
---



## Setup

Load the necessary library


```{r libraries, warning=FALSE, message=FALSE}
library(dplyr)
```

## Read all the files


```{r loadData}
mutation_data = read.table("data/original/mc3.v0.2.8.PUBLIC.maf.gene_vclass_HGVSp_sample_likelyDriverLoose.tsv", head = TRUE)

patient_data = read.csv("data/original/clinical_PANCAN_patient_with_followup.tsv", header = TRUE, sep = "\t")

cnv_data = read.csv("data/original/all_thresholded.by_genes_whitelisted.tsv", header = TRUE, sep = "\t")

cnv_data_preprocessed <- cnv_data

names(cnv_data_preprocessed)[1] ="Hugo_Symbol"
```

## Preprocess the data

Add the tumor type information from patient_data to mutation_data


```{r readData}
mutation_data_merged <- merge(mutation_data, patient_data, by="bcr_patient_barcode", all.x = TRUE, all.y = FALSE) %>%
  select(bcr_patient_barcode, acronym, Hugo_Symbol, Variant_Classification)
```

```{r addTumorType}
mutation_data_merged$bcr_patient_barcode <- gsub(pattern = "-", replacement= ".", mutation_data_merged$bcr_patient_barcode)
```

Rename the "Variant_Classification" column to "Mutation"

```{r rename}
mutation_data_merged_simplified <- mutation_data_merged
colnames(mutation_data_merged_simplified)[colnames(mutation_data_merged_simplified) == "Variant_Classification"] <- "Mutation"
```

Modify the "Mutation" column in the mutation_data_merged_simplified table, changing any non-NA value to 1

```{r modify}
mutation_data_merged_simplified$Mutation[!is.na(mutation_data_merged_simplified$Mutation)] <- 1
```


Shorten and edit the names in cnv_data so that they can be matched with patient_data
 

```{r editNames}
names(cnv_data_preprocessed) <- substring(names(cnv_data_preprocessed), 1, 12)

names(cnv_data_preprocessed) <- gsub(pattern = "[.]", replacement= "-", names(cnv_data_preprocessed))
```

Copy column names to first row, so that it will be easier to work on it (colnames should be unique)


```{r copyColnames}
cnv_data_preprocessed <- rbind(colnames(cnv_data_preprocessed), cnv_data_preprocessed)
```

Find acronyms corresponding to the patient barcodes and overwrite them


```{r findAcronyms}
patient_barcodes <- unique(patient_data$bcr_patient_barcode)

for (colnum in 4:ncol(cnv_data_preprocessed)) {
  if (cnv_data_preprocessed[1, colnum] %in% patient_barcodes) {
    cnv_data_preprocessed[1, colnum] = patient_data$acronym[which(cnv_data_preprocessed[1, colnum] == patient_data$bcr_patient_barcode)]
  } else {
    cnv_data_preprocessed[1, colnum] = NA
  }
}
```

Remove columns that did not have matching acronyms


```{r removeCol1}
cnv_data_preprocessed <- cnv_data_preprocessed[ , colSums(is.na(cnv_data_preprocessed[1,]))==0]
```

Remove two unnecesary columns


```{r removeCol2}
cnv_data_preprocessed <- cnv_data_preprocessed[, -c(2:3)]
```

Change the colnames to match the ones in mutations data


```{r changeColnames}
colnames(cnv_data_preprocessed) <- gsub(pattern = "-", replacement= ".", colnames(cnv_data_preprocessed))
```

Create list with all acronyms in the right order


```{r acronymsList}
acronyms <- as.character(as.vector(cnv_data_preprocessed[1,]))

acronyms <- acronyms[-1]
```

Remove the first row


```{r remoweRow}
cnv_data_preprocessed <- cnv_data_preprocessed[-1, ]
```

## Save the files

Save merged mutation data


```{r saveMerged}
dir.create("data/preprocessed", recursive = TRUE)

write.csv(mutation_data_merged, "data/preprocessed/mutation_data_merged.csv", row.names = FALSE)

write.csv(mutation_data_merged_simplified, "data/preprocessed/mutation_data_merged_simplified.csv", row.names = FALSE)
```


Save preprocessed cnv data


```{r savePreprocessed}
write.csv(cnv_data_preprocessed, "data/preprocessed/cnv_data_preprocessed.csv")
```

Save all acronyms


```{r saveAcronyms}
saveRDS(acronyms, file = "data/preprocessed/acronyms.rds")
```


